\protect Connecting every bit of knowledge: \\ The structure of Wikipedia's First Link Network

### 2016-09-07 writeup/2016-04wikipedia-network/wikipedia-network.body.tex
traversal funnels v centrality measures

@@

While traversal visits measure accumulation, each article's first link also 
influences the shape of the FLN. 
At a point of accumulation, a single article's first link 
can exert great influence over the shape of the FLN by directing many
references on a particular path. To distinguish between an article 
that simply happened to fall within a cycle from an article funneling 
many first links, we develop a second metric called {\it traversal funnels}.
To measure traversal funnels, we traverse the FLN in the same manner as we 
did for traversal visits, but end a path once we enter a cycle.
We are then able to distinguish between an article related to many other ideas
only by virtue of its place in a cycle, from an article exerting influence over where the first links flow. 
In our simple network 
(see Fig.~\ref{fig:Traversal Funnels}), article C 
directs the flow of links towards the 3-cycle, while articles A and B are 
recipients of the flow---without exerting direct influence themselves.

By considering the FLN not only as collection of directly linked pairs of articles, but
as a flow, with accumulated references, cycles, and path-connected groups, we can explore how the many articles in Wikipedia are organized and related.

>>>>

\subsection{Traversal Visits}
\label{Traversal Visits}

@@

\begin{figure}[tp!]
  \includegraphics[width=\columnwidth]{centrality_measures.pdf}
  \caption{
    \textbf{Centrality Measures}
    The table above compares standard centrality measures on the sample network against traversal funnels.
    }
  \label{fig:Centrality Measures}

While traversal visits measure accumulation, each article's first link also 
influences the shape of the FLN. 
At a point of accumulation, a single article's first link 
can exert great influence over the shape of the FLN by directing many
references on a particular path. To distinguish between an article 
that simply happened to fall within a cycle from an article funneling 
many first links, we develop a second metric called {\it traversal funnels}.
To measure traversal funnels, we traverse the FLN in the same manner as we 
did for traversal visits, but end a path once we enter a cycle.
We are then able to distinguish between an article related to many other ideas
only by virtue of its place in a cycle, from an article exerting influence over where the first links flow. 
In the sample network 
(see Fig.~\ref{fig:Traversal Funnels}), article C 
directs the flow of links towards the 3-cycle, while articles A and B are 
recipients of the flow---without exerting direct influence themselves.

To compute traversal funnels on a directed network, begin at each node: 1. Traverse the edge to the next node 2. Increment the next node's count 3. Repeat while the next node is unvisited. The resulting count is each node's number of traversal funnels: the number of paths the node directs up to the start of a cycle.

As a measure of influence, traversal funnels differs from traditional centrality measures such as degree centrality, betweenness centrality, and eigenvector centrality. 
Degree centrality characterizes how many articles have a first link to a particular topic, restricting influence to a single link. 
Betweenness centrality and eigenvalue centrality capture flow within the network more than a single link away. Unlike traversal funnels however, betweenness centrality and eigenvalue centrality reward nodes that simply happen to be in a cycle. In our sample network (see Fig.~\ref{fig:Centrality Measures}), node A is relatively central using both betweenness and eigenvector centrality although node A is simply the recipient of paths directed by node C. Eigenvector centrality even assigns the same centrality to nodes A, B, and C. In contrast, traversal funnels defines a path without a predetermined end node, following the progression of links to its culmination at the start of a cycle. Furthermore, traversal funnels provides an interpretable metric: a node's traversal funnels is the number of paths the node directs to the start of a cycle. 
Consequently, traversal funnels allow us to isolate the paths directed by node C without any spill over effect on other nodes in its cycle.

In addition, traversal funnels extends to directed networks with more than one outward edge. In such a network, there may be multiple paths associated with a single starting node. 
The same interpretation for traversal funnels as the number of paths a node directs to a cycle still holds. 
In contrast to betweenness centrality, traversal funnels would not favor the shortest path over any other. Traversal funnels equally weighs all paths up to a cycle.

For our exploration of the FLN, by considering the FLN not only as collection of directly linked pairs of articles, but
as a flow, with accumulated references, cycles, and path-connected groups, we can explore how the many articles in Wikipedia are organized and related.
